---
title: "R Markdown Example"
author: "Sean Horvath"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
subtitle: Solar Radiation
---

See appendix for source code.
```{r echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE)
```
```{r message=FALSE}
# Load libraries:
library(knitr);library(reshape2);library(dplyr);library(ggplot2)
# Create function to calculate all TOA variables:
toa_radiation <- function(year,month,day,hour,lat){
  return_dataframe <- data.frame(declination=NA,
  Z=NA,
  Hs=NA,
  r=NA,
  Seff=NA,
  I_day=NA,
  I=NA)
  # Declination
  day_of_interest <- as.Date(paste0(year,'-',month,'-',day))
  reference_day <- as.Date(paste0(year,'-06-21'))
  n <- day_of_interest-reference_day
  if(n<0){
  reference_day <- as.Date(paste0(year-1,'-06-21'))
  n <- day_of_interest-reference_day
  }
  return_dataframe$declination <- 23.5*cos(0.986*as.numeric(n)*pi/180)
  # Z
  lat <- lat*pi/180
  return_dataframe$Z <- acos(sin(lat)*sin(return_dataframe$declination*pi/180)+cos(lat)*cos(return_dataframe$declination*pi/180)*cos((15*(hour-12))*pi/180))*180/pi
  # Hs
  temp <- -tan(lat)*tan(return_dataframe$declination*pi/180)
  if(temp>1){
    temp=1
  } else if(temp<(-1)){
      temp=-1
  }
  Hs_value <- acos(temp)*180/pi
  return_dataframe$Hs <- Hs_value
  # r
  day_of_interest <- as.Date(paste0(year,'-',month,'-',day))
  reference_day <- as.Date(paste0(year,'-07-04'))
  n <- day_of_interest-reference_day
  if(n<0){
  reference_day <- as.Date(paste0(year-1,'-07-04'))
  n <- day_of_interest-reference_day
  }
  r_value <- 149*10^6+2.5*10^6*cos(0.986*as.numeric(n)*pi/180)
  return_dataframe$r <- r_value
  # Seff
  return_dataframe$Seff <- (149*10^6/return_dataframe$r)^2*1367
  # I_day
  return_dataframe$I_day <- 435*(149*10^6/return_dataframe$r)^2*(return_dataframe$Hs*pi/180*sin(lat)*sin(return_dataframe$declination*pi/180)+cos(lat)*cos(return_dataframe$declination*pi/180)*sin(return_dataframe$Hs*pi/180))
  # I
  return_dataframe$I <- 1367*(149000000/return_dataframe$r)^2*cos(return_dataframe$Z*pi/180)
  return(return_dataframe)
}
```

**Problem 1**  
**a)** Code Output:
```{r}
# Problem 1a)
doy <- data.frame(month=c(rep(1,31),rep(2,28),rep(3,31),rep(4,30),
                            rep(5,31),rep(6,30),rep(7,31),rep(8,31),
                            rep(9,30),rep(10,31),rep(11,30),rep(12,31)),
                    day=c(seq(1,31),seq(1,28),seq(1,31),seq(1,30),
                          seq(1,31),seq(1,30),seq(1,31),seq(1,31),
                          seq(1,30),seq(1,31),seq(1,30),seq(1,31)),
                    Seff=rep(NA,365))
  for(i in 1:dim(doy)[1]){
    doy$Seff[i] <- toa_radiation(2001,doy$month[i],doy$day[i],12,40)$Seff
  }
  print(range(doy$Seff))
  print(range(doy$Seff)/1367)
```
The range of the effective solar constant is 1322-1414 $\frac{W}{m^{2}}$  
Or as a percentage of the true solar constant: 0.97%-1.03%  

**b)** Code Output:
```{r}
# Problem 1b)
print(paste0('Maximum: ',format(as.Date(paste0('2001-',
                                             doy$month[which.max(doy$Seff)],'-',
                                               doy$day[which.max(doy$Seff)])),
                                "%B %d")))
print(paste0('Minimum: ',format(as.Date(paste0('2001-',
                                               doy$month[which.min(doy$Seff)],
                                            '-',doy$day[which.min(doy$Seff)])),
                                "%B %d"))) 
```
The dates that maximum and minimum occur are January 3 and July 4, repectively.  

**Problem 2**  
Values in Degrees
```{r}
# Problem 2)
Z_dates <- c('12-21','02-05','03-21','06-21')
Z_lats <- c(21,40,60,71)
Z_df <- expand.grid(date=Z_dates,lat=Z_lats)

for(i in 1:dim(Z_df)[1]){
  Zmonth <- as.numeric(substr(Z_df$date[i],1,2))
  Zday <- as.numeric(substr(Z_df$date[i],4,5))
  Zlat <- Z_df$lat[i]
  Z_df$Z[i] <- toa_radiation(2018,Zmonth,Zday,12,Zlat)$Z
}
Z_df <- dcast(Z_df,lat~date,value.var='Z')
Z_df <- Z_df[,-1]
rownames(Z_df) <- c('Honolulu, HI (21 N)',
                    'Boulder, CO (40 N)',
                    'Oslo, Norway (60 N)',
                    'Barrow, AK (71 N)')
colnames(Z_df) <- c('Dec. 21','Feb. 5','Mar. 21','June 21')
knitr::kable(Z_df)
```

**Problem 3**  
Values in $\frac{W}{m^{2}}$
```{r}
# Problem 3)
I_dates <- c('12-21','02-05','03-21','06-21')
I_lats <- c(21,40,60,71)
I_df <- expand.grid(date=I_dates,lat=I_lats)

for(i in 1:dim(I_df)[1]){
  Imonth <- as.numeric(substr(I_df$date[i],1,2))
  Iday <- as.numeric(substr(I_df$date[i],4,5))
  Ilat <- I_df$lat[i]
  I_df$I[i] <- toa_radiation(2018,Imonth,Iday,12,Ilat)$I
}
I_df <- dcast(I_df,lat~date,value.var='I')
I_df <- I_df[,-1]
rownames(I_df) <- c('Honolulu, HI (21 N)',
                    'Boulder, CO (40 N)',
                    'Oslo, Norway (60 N)',
                    'Barrow, AK (71 N)')
colnames(I_df) <- c('Dec. 21','Feb. 5','Mar. 21','June 21')
kable(I_df)
```

**Problem 4**  
Day length in hours.  
Daily averaged solar insolation in $\frac{W}{m^{2}}$
```{r}
# Problem 4)
I_dates <- c('12-21','02-05','03-21','06-21')
I_lats <- c(21,40,60,71)
I_df <- expand.grid(date=I_dates,lat=I_lats)

for(i in 1:dim(I_df)[1]){
  Imonth <- as.numeric(substr(I_df$date[i],1,2))
  Iday <- as.numeric(substr(I_df$date[i],4,5))
  Ilat <- I_df$lat[i]
  I_df$Hs[i] <- toa_radiation(2002,Imonth,Iday,12,Ilat)$Hs
  I_df$I_day[i] <- toa_radiation(2002,Imonth,Iday,12,Ilat)$I_day
}

I_df$Hs <- I_df$Hs*2/15

Hs_df <- dcast(I_df,lat~date,value.var='Hs')
colnames(Hs_df) <- c('lat','Dec. 21','Feb. 5','Mar. 21','June 21')

Iday_df <- dcast(I_df,lat~date,value.var='I_day')
colnames(Iday_df) <- c('lat','Dec. 21','Feb. 5','Mar. 21','June 21')

combined_df <- bind_rows(Hs_df,Iday_df) %>%
  arrange(lat)
combined_df <- combined_df[,-1]
rownames(combined_df) <- c('Honolulu, HI (21 N) day length',
                           'Honolulu, HI (21 N) daily avg solar',
                           'Boulder, CO (40 N) day length',
                           'Boulder, CO (40 N) daily avg solar',
                           'Oslo, Norway (60 N) day length',
                           'Oslo, Norway (60 N) daily avg solar',
                           'Barrow, AK (71 N) day length',
                           'Barrow, AK (71 N) daily avg solar')
kable(combined_df)
```

**Problem 5**  
```{r}
# Problem 5)
diurnal_I <- data.frame(Hour=rep(seq(0,24),2),
                        Insolation=rep(NA,50),
                        City=rep(c('Boulder','Barrow'),each=25))
for(i in 1:dim(I_df)[1]){
  for(j in 1:25){
    diurnal_I$Insolation[j] <- toa_radiation(2002,6,21,
                                            diurnal_I$Hour[j],40)$I
    diurnal_I$Insolation[j+25] <- toa_radiation(2002,6,21,
                                            diurnal_I$Hour[j],71)$I
  }
}
diurnal_I$Insolation[which(diurnal_I$Insolation<0)] <- 0

ggplot(diurnal_I,aes(x=Hour,y=Insolation,color=City)) +
  geom_line(aes(group=City)) +
  ylab(expression(paste('Hourly Solar Insolation, ',over(W,m^2)))) +
  ggtitle('Diurnal Cycle of Incoming Solar Radiation')
```

Latitude is the culprit here, driving the differences in diurnal cycles of these two locations. Because the latitude of Barrow is above 66.5$^\circ$ it experiences 24-hours of sunlight during parts of the summer (such as on June 21). This is why the radiation flux never drops to 0. Boulder on the other hand still experiences night, so there are times in the early morning and late night when the radiation flux becomes 0.  

The intensity of radiation is greater at noon in Boulder because of latitude as well. On June 21 the sun is directly overhead at noon at 23.5$^\circ$ north. Because Boulder's latitude is closer to that than Barrow, radiation from the sun is coming in at less of an angle, resulting in the higher flux. Finally, the slope of radiation flux is smaller for Barrow because the angle of the sun doesn't change as much throughout the day compared to Boulder.  


#### Code Appendix
```{r code = readLines(knitr::purl("C:/Users/seanm/Documents/R Tutorial/Example.Rmd", documentation = 1)), echo = T, eval = F}
```




